Is the Future of Authenticity In Our Heads?\\\medskip \large Moving Passthoughts From the Lab to the World

### 2017-07-03 authors.tex
Minor notes

\author{Nick Merrill, Max T Curran, John Chuang}
\affiliation{%
  \institution{BioSENSE, UC Berkeley School of Information}
  \city{Berkeley} 
  \state{California, USA} 
}
\email{ffff, mtcurran, john.chuang @berkeley.edu}

@@

\begin{abstract}
\textit{Passthoughts authentication}, in which a user thinks a secret thought to log in to services or devices,
provides two factors of authentication (knowledge and inherence) in a single step.
Since its proposal in 2005, passthoughts enjoyed a number of successful empirical studies.
In this paper, we renew the promise of passthoughts authentication,
reviewing the main challenges that passthoughts must overcome in order to move from the lab to the real world.
We propose two studies,
which seek different angles at the fundamental questions we pose.
We conclude with future oppportunities, and challenges, for passthoughts' broader deployment as a tool for authentication.
In doing so, we raise novel possibilities for authentication broadly,
such as ``organic passwords'' that change naturally over time,
or systems that reject users who are not acting quite ``like themselves.''
\end{abstract}

>>>>

\author{Nick Merrill, Max T Curran, John Chuang}
\affiliation{%
  \institution{BioSENSE, UC Berkeley School of Information}
  \city{Berkeley, California} 
  \state{USA} 
}
\email{email}

@@

\begin{abstract}
Passthoughts, in which a user thinks a secret thought to log in to services or devices, provides two factors of authentication (knowledge and inherence) in a single step. Since its proposal in 2005, passthoughts enjoyed a number of successful empirical studies. In this paper, we renew the promise of passthoughts authentication, reviewing the main challenges that passthoughts must overcome in order to move from the lab to the real world. We propose two studies, which seek different angles at the fundamental questions we pose. We conclude with future opportunities, and challenges, for passthoughts’ broader deployment as a tool for authentication. In doing so, we raise novel possibilities for authentication broadly, such as ``organic passwords'' that change naturally over time, or systems that reject users who are not acting quite ``like themselves.''
\end{abstract}

%
% Here's a visual tool for generating the CCS categories LaTeX below:
% 
%    http://dl.acm.org/ccs/ccs.cfm
%
% Use it to generate the two b
%

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003120.10003121.10011748</concept_id>
<concept_desc>Human-centered computing~Empirical studies in HCI</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[300]{Human-centered computing~Empirical studies in HCI}

### 2017-07-03 paper.tex
Minor notes

% Created 2017-04-14 Fri 12:43
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 25.1.1 (Org mode 9.0.4)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:org3fa1a74}

@@

This paper motivates and contextualizes two studies that could begin to address these challenges (Section 5).
While these studies attack two sides of passthoughts' chances as a real-world technology
this study serves not just to propose solutions, but to raise questions about what authentication could and should be able to do.
While passthoughts is not without risks (Section 6),
it raises the possiblity for several interesting paths,
including exploration of systems that could enable novel types of authentication (Section 7).

\section{Background}
\label{sec:orgdb93e8e}

@@

\subsection{One-step, multi-factor authentication}
\label{sec:orgd0ea216}

@@

Other work has attempted to provide multiple factors of authentication in one step.
Some work has tested behavioral authentication methods such as keystroke dynamics, or voice. In both cases, the knowledge factor (password or passphrase) and
inherence factor (typing rhythm or speaker's voice) are employed \cite{Monrose1997}.
In contrast, the Nymi band supports one-step two-factor authentication via the inherence
factor (cardiac rhythm that is supposed to be unique to each individual) and the
possession factor (the wearing of the band on the wrist) \cite{Nymi}.
Custom-built EEG devices could incorporate an added possession factor 
to the already two-step authentication provided by passthoughts \cite{Curran2017}.

Authenticators are susceptible to a variety of attacks. 
One common attack on passwords is ``shoulder surfing,'' in which an attack uses visual or other cues to steal, or improve the chances of guessing, a target's chosen secret. 
Passthoughts mitigates this attack by nature of the mental gesture:
since the expression of a passthought is not externally visible, the authenticator is impervious to shoulder surfing attacks.

\subsection{Passthought authentication}
\label{sec:orgc70a625}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:orgf59991b}

@@

\section{User attitudes and perceptions}
\label{sec:org89c8777}

@@

\subsection{Contending with mind-reading machines}
\label{sec:org6675840}

@@

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:org86d6f6a}

The survey we report on here, currently in-progress, examines how people's beliefs differ given device ownership, and their membership in one of two groups: Mechanical Turk workers, or people enrolled in Health-e-Heart, a massive (n > 40,000), longitudinal study, in which volunteers fill out surveys about themselves, and/or upload data from biomedical self-tracking devices, over the course of several years \cite{Estrin2010a}.
In one portion of the survey, we ask subjects to rate a number of different biosensors in order of how likely individual's believe each sensor is to reveal what ``a person is thinking or feeling'' (Figure \ref{fig:rank}).

@@

\section{Diversity and security of passthoughts}
\label{sec:orgc9f9f42}

@@

\section{Two studies on passthoughts}
\label{sec:orgd8698f9}

@@

\subsection{A real-time passthought authenticator}
\label{sec:orgcba969d}

@@

\subsubsection{Study protocol}
\label{sec:org2ee4043}

@@

\subsubsection{The effect of feedback}
\label{sec:org17389d3}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:org953b2b7}

In addition to these findings, the data generated during this study could help test 
a third hypothesis: that the continual re-training of passthought classifiers might help boost classification performance over time,
especially in the face of shifting signals.
Offline, we can train each classifier, for each subject, to achieve its post-calibration state.
Next, we can run each reading recorded from a particular participant through the trained classifier.
If the classifier accepts the reading, we can then re-train the classifier, 
adding this reading to the corpus of positive examples.
In a separate, \emph{negative calibration} condition, 
we also re-train the classifier with rejected readings as negative examples.

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:org8725ec2}

@@

\subsubsection{Study protocol}
\label{sec:orgcf929e2}
A small group of subjects could wear a working, recording EEG device, whether or not it provides feedback, in a variety of settings for some number of days,
having subjects journal their experiences and asking them specifically what they feel someone might be able to know about them from the EEG signals they record.
At the same time, we could use this study as an opportunity to collect a much larger, and more diverse corpus.
To aid in the collection of signals that are specific to our problem of passthought authentication,
subjects in this study might be prompted to perform a variety of tasks at a few checkpoints throughout the day.
With the data collected during this study, we could easily simulate passthought accuracy on a much more realistic (and representative) sample of readings.

@@

\subsubsection{A more diverse corpus}
\label{sec:org5956bb5}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:org35d68ac}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how likely or unlikely we are to observe a given passthought in context of a particular person's neural signals
Such analysis between subjects could help shed light how given passthoughts are expressed uniquely between individuals.

@@

\subsubsection{Usability and attitudes}
\label{sec:orgae2a266}

@@

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:org4abf94a}

@@

\subsection{Privacy}
\label{sec:orga7fe0ca}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:orga3eb0b3}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

@@

\subsection{Tradeoffs between security and privacy}
\label{sec:orgf5901f8}

@@

\section{Further Future Directions}
\label{sec:orgce82d31}

This paper so far has motivated two future studies on passthoughts,
and discussed potential risks intrinsic to the development of passthoughts systems.
With these risks in mind, the present section explores some of the exciting possibilities that could unfold after the immediate priorities described in the prior sections.

\subsection{Continuous authentication}
\label{sec:org6cddcf6}

@@

\subsection{Organic passwords}
\label{sec:orge8a2dcd}

If EEG signals are nonstationary (changing over time), passthoughts will require continuous re-calibration to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as voice or keystroke dynamics \cite{Monrose1997}, which may change more slowly for individuals, if they change at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Authentication and the self}
\label{sec:org88a2cde}

@@

Such a scenario raises serious ethical, legal, and even philosophical questions. 
How does such a system conform to accepted definitions of a ``person''?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as we are aware, novel, 
and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this developing method for authentication.
\subsection{Passthoughts by any other sensor?}
\label{sec:org6827b98}

@@

\subsection{Health, neuroscience and BCIs}
\label{sec:orga3c747f}

@@

\section{Conclusion}
\label{sec:orgeeb8298}

>>>>

% Created 2017-07-03 Mon 10:49
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 25.1.1 (Org mode 9.0.4)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:orgf1c0eda}

@@

In addition, computing broadly has undergone many sea changes since passthought's initial proposal in 2005.
Biosensing (the sensing of humans) has proliferated throughout our everyday life, due both to the sensors ubiquitous in smartphones, and to general improvements to infrastructures of data collection and machine learning at scale. 
Do these changes make passthoughts less appealing compared to other authentication strategies that these "sea changes" have made possible?

This paper aims both to re-validate passthought's promise as an authentication paradigm (Section 2), and to 
motivate and contextualize two studies that could begin to address the challenges posed above (Section 5).
While passthoughts is not without risks (Section 6),
it could lead to novel \emph{types} of authentication,
which challenge (or at least destabilize) dominant assumptions about what authentication 
can and should be able to do. (Section 7).

\section{Background}
\label{sec:org6b9dd4a}

@@

@@

\subsection{One-step, multi-factor authentication}
\label{sec:org526fbba}

Since passthought's initial proposal in 2005,
more ubiquitous sensing and computing has enabled a number of other strategies for achieving two factors of authentication in a single step. 
Some work has tested behavioral authentication methods such as keystroke dynamics, or voice. In both cases, the knowledge factor (password or passphrase) and
inherence factor (typing rhythm or speaker's voice) are employed \cite{Monrose1997}.
In contrast, the Nymi band supports one-step two-factor authentication via the inherence
factor (cardiac rhythm that is supposed to be unique to each individual) and the
possession factor (the wearing of the band on the wrist) \cite{Nymi}.
More recent attempts have also used gait (from cellphone accelorometers) to perform authentication \cite{UnifyID2017}.

However, these existing strategies are susceptible to a variety of attacks. 
Nymi, for example, does not have a knowledge factor, making it impossible for the user to change the authentication token if the device and biosignal have been compromised.
Keystroke dynamics, voice, and gait are all susceptable to ``shoulder surfing,'' in which an attack uses visual or other cues to steal, or improve the chances of guessing, a target's chosen secret. 
Passthoughts mitigates this attack by nature of the mental gesture.
Since the expression of a passthought is not externally visible, the authenticator is impervious to shoulder surfing attacks.
Since the thought a chosen secret (knowledge factor), it can be changed if compromised.

\subsection{Passthought authentication}
\label{sec:org6b3c69e}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:org0f217d8}

@@

\section{User attitudes and perceptions}
\label{sec:org0009958}

@@

\subsection{Contending with mind-reading machines}
\label{sec:org5b8dcbd}

@@

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:orga19eaac}

The survey we report on here, currently in-progress, examines how people's beliefs differ given device ownership, and their membership in one of two groups: Mechanical Turk workers, or people enrolled in Health-e-Heart, a massive (n > 40,000), longitudinal study, in which volunteers fill out surveys about themselves, and/or upload data from biomedical self-tracking devices, over the course of several years \cite{Estrin2010a}.
In one portion of the survey, we ask subjects to rate a number of different biosensors in order of how likely individual's believe each sensor is to reveal what ``a person is thinking or feeling'' (Figure \ref{fig:rank}).
This section reports on a subset of Mechanical Turk workers (n=100) and Health-e-Hearth subjects (n=100).

@@

\section{Diversity and security of passthoughts}
\label{sec:org4fff047}

@@

\section{Two studies on passthoughts}
\label{sec:org9bd76c2}

@@

\subsection{A real-time passthought authenticator}
\label{sec:org65e0cf5}

@@

\subsubsection{Study protocol}
\label{sec:org0bd4853}

@@

\subsubsection{The effect of feedback}
\label{sec:org4f731b0}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:orgfa27859}

In addition to these findings, the data generated during this study could help test 
a third hypothesis: that the continual re-training of passthought classifiers might help boost classification performance over time,
especially in the face of shifting signals.
Offline, we can train each classifier, for each subject, to achieve its post-calibration state.
Next, we can run each reading recorded from a particular participant through the trained classifier.
If the classifier accepts the reading, we can then re-train the classifier, 
adding this reading to the corpus of positive examples.
In a separate, \emph{negative calibration} condition, 
we can also re-train the classifier with rejected readings as negative examples.
This condition should reduce false acceptances from the target subject, re-inforcing our authenticator's knowledge factor.

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:org7ef7168}

@@

\subsubsection{Study protocol}
\label{sec:org7967cd6}

A small group of subjects could wear a working, recording EEG device, whether or not it provides feedback, in a variety of settings for some number of days,
having subjects journal their experiences and asking them specifically what they feel someone might be able to know about them from the EEG signals they record.
At the same time, we could use this study as an opportunity to collect a much larger, and more diverse corpus.
To aid in the collection of signals that are specific to our problem of passthought authentication,
subjects in this study might be prompted to perform a variety of tasks at a few checkpoints throughout the day.
With the data collected during this study, we could easily simulate passthought accuracy on a much more realistic (and representative) sample of readings.

@@

\subsubsection{A more diverse corpus}
\label{sec:org6c1e1e9}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:org9ef3205}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how likely or unlikely we are to observe a given passthought in context of a particular person's neural signals
Such analysis between subjects could help shed light how given passthoughts are expressed uniquely between individuals.

@@

\subsubsection{Usability and attitudes}
\label{sec:orge398ca0}

@@

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:orgbb7cf43}

@@

\subsection{Privacy}
\label{sec:orga1e2eb9}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:org759dea8}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

@@

\subsection{Tradeoffs between security and privacy}
\label{sec:orgd71b4f9}

@@

\section{Further Future Directions}
\label{sec:org9fefd9f}

This paper so far has motivated two future studies on passthoughts,
and discussed potential risks intrinsic to the development of passthoughts systems.
With these risks in mind, the present section explores some of the exciting possibilities that could unfold after the immediate priorities described in the prior sections.
Through the lens of passthoughts, we hope to use this discussion as an opportunity to challenge (or at least destabilize) dominant assumptions in authentication.

\subsection{Continuous authentication}
\label{sec:org40ff196}

@@

\subsection{Organic passwords}
\label{sec:org74d0a56}

If EEG signals are nonstationary (changing over time), passthoughts will require continuous re-calibration to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as voice or keystroke dynamics \cite{Monrose1997}, which may change more slowly for individuals, if they change at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Authentication and the self}
\label{sec:org740e10b}

@@

Such a scenario raises serious ethical, legal, and even philosophical questions. 
How does such a system conform to accepted definitions of a ``person''?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as we are aware, novel, 
and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this developing method for authentication.
\subsection{Passthoughts by any other sensor?}
\label{sec:org8dc3b22}

@@

\subsection{Health, neuroscience and BCIs}
\label{sec:org037222f}

@@

\section{Conclusion}
\label{sec:orgc0cdb64}
